---
title: "Car Dataset Analysis"
author: "Deepika Ganesan"
format: html
editor: visual
css: "styles.css"
---

## Data Summary Report

This report provides an analysis of the \`mtcars\` dataset, including tables and charts.

## Data Summary Table

The table below shows summary statistics (mean and standard deviation) for each variable in the dataset:

```{r include=FALSE}
# Load necessary library
library(dplyr)

# Rename columns 
mtcars_renamed <- mtcars %>%
  rename(
    Miles_per_Gallon = mpg,
    Cylinders = cyl,
    Displacement = disp,
    Horsepower = hp,
    Rear_Axle_Ratio = drat,
    Weight = wt,
    Quarter_Mile_Time = qsec,
    Engine_Shape = vs,
    Transmission = am,
    Gears = gear,
    Carburetors = carb
  )

# Generate summary statistics
summary_table <- mtcars_renamed %>%
  summarise(across(everything(), list(mean = mean, sd = sd)))
```

```{r echo=FALSE}
# Display table using knitr
knitr::kable(summary_table, format = "html", table.attr = "class='styled-table'")
```

## Charts

### Scatter Plot

```{r echo=FALSE}
# Load necessary library
library(ggplot2)

# Generate the plot
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(color = "#FF6F61", size = 3) +
  geom_smooth(method = "lm", color = "#1B9E77", se = FALSE) +
  theme_minimal() +
  labs(title = "Miles Per Gallion vs Weight",
       x = "Weight (1000 lbs)",
       y = "Miles per Gallon") +
  theme(
    plot.title = element_text(size = 18, face = "bold", color = "#FF6F61", hjust = 0.5),
    axis.title = element_text(size = 14, color = "#333"),
    axis.text = element_text(size = 12, color = "#333"),
    axis.ticks = element_blank(),
    panel.grid.major = element_line(color = "#ddd"),
    panel.grid.minor = element_blank()
  )
```

### Bar Chart

```{r echo=FALSE}
# Generate the bar chart: Cylinders vs. Miles per Gallon (MPG)
ggplot(mtcars_renamed, aes(x = factor(Cylinders), y = Miles_per_Gallon, fill = factor(Cylinders))) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge", show.legend = FALSE) +
  scale_fill_manual(values = c("#FF6F61", "#FFB84D", "#00A5A7", "#FFD700", "#1B9E77")) +
  theme_minimal() +
  labs(
    title = "Average Miles per Gallon vs Number of Cylinders",
    x = "Number of Cylinders",
    y = "Average Miles per Gallon (MPG)"
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold", color = "#FF6F61", hjust = 0.5),
    axis.title = element_text(size = 14, color = "#333"),
    axis.text = element_text(size = 12, color = "#333"),
    axis.ticks = element_blank(),
    panel.grid.major = element_line(color = "#ddd"),
    panel.grid.minor = element_blank()
  )
```

## Insights

### Scatter Plot

-   Negative correlation between weight and fuel efficiency.

-   Lighter vehicles are more fuel-efficient.

-   Other factors also influence fuel efficiency.

### Bar Chart

-   More cylinders, lower fuel efficiency.

-   4-cylinder engines are most fuel-efficient.

-   8-cylinder engines are least fuel-efficient.
